---
title: "Ripped Foil - Children"
author: "Zeynep Civelek"
date: "June 10, 2020"
output: word_document
---

```{r, include=FALSE}
#This sets general options for the code chunks in the R Markdown file. The echo, warning, and message = FALSE hide the code chunks, warning messages, and other messages, where the ‘include=true’ will make all figures appear in the text. You can set some of these variables to TRUE, and hit Knit to see what they change. 
```

```{r, include=FALSE}
#PREPARE
R.Version()#for referencing, shows you which R version you are using
rm(list=ls())#removes any other items in your workspace
ls()#check whether workspace is empty
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Directory is set, data is added and all the libraries necessary for running the analyses are loaded.

#setwd("C:/Users/Zeynep/Desktop/ripped-foil/Ripped Foil")
ripchild<-read.csv("ripped_child.csv", header=T)
#all the libraries necessary to run the analyses
library(lme4)
library(readr)
library(tidyverse)
library(sjPlot)
#install.packages('TMB', type='source')
library(ggthemes)
library(gridExtra)
library(reshape2)
library(car)
#library(psych)
library("ggpubr")
library(dplyr)

```
##Exploring the data

The structure of the data is shown below with all the variable names and levels of measurement.

```{r echo=FALSE}
ripchild$age <- as.factor(ripchild$age)# makes continuous variables categorical
ripchild$id<-as.factor(ripchild$id)
ripchild$sex<-as.factor(ripchild$sex)
str(ripchild)
summary(ripchild)
sd(ripchild$agemonths)
sd(ripchild$trialno)
```

```{r}
###Getting the age, sex distribution table
ripchild_penfirst <- ripchild %>%
  filter(order=="penfirst") %>%
  group_by(id, sex, agemonths) %>% 
  summarize(correct = mean(correct))
boys_pen <- subset(ripchild_penfirst, sex == "2")
mean(boys_pen$agemonths)
sd(boys_pen$agemonths)
min(boys_pen$agemonths)
max(boys_pen$agemonths)
girls_pen <- subset(ripchild_penfirst, sex == "1")
mean(girls_pen$agemonths)
sd(girls_pen$agemonths)
min(girls_pen$agemonths)
max(girls_pen$agemonths)


ripchild_stickerfirst <- ripchild %>%
  filter(order=="stickerfirst") %>%
  group_by(id, sex, agemonths) %>% 
  summarize(correct = mean(correct))
boys_sticker <- subset(ripchild_stickerfirst, sex == "2")
mean(boys_sticker$agemonths)
sd(boys_sticker$agemonths)
min(boys_sticker$agemonths)
max(boys_sticker$agemonths)
girls_sticker <- subset(ripchild_stickerfirst, sex == "1")
mean(girls_sticker$agemonths)
sd(girls_sticker$agemonths)
min(girls_sticker$agemonths)
max(girls_sticker$agemonths)
```



We decided to use age as a continuous variable (age in months) in our main analyses rather than categorical to have more power.The distribution of age:

```{r pressure, echo=FALSE}
hist(ripchild$agemonths, xlab="Age in months (36-72)", ylab="Frequency")

```

Next I aggregated the trial-by-trial data to create one score per child for Test and Transfer phases as well as an overall score so I can check the distributions of scores.

```{r echo=FALSE}
ripchild_individual <- ripchild %>%
  group_by(id, age, agemonths) %>% 
  summarize(correct = mean(correct))%>%
  ungroup()%>%
  group_by(agemonths)%>%
    add_count(correct)

ripchild_test <- ripchild %>%
  filter(order3=="test") %>%
  group_by(id, agemonths) %>% 
  summarize(correct = mean(correct))

ripchild_transfer <- ripchild %>%
  filter(order3=="transfer") %>%
  group_by(id, agemonths) %>% 
  summarize(correct = mean(correct)) 

```

##Histograms for Test and Transfer Phases and an overall score (ripchild_individual)

```{r echo=FALSE}
hist(ripchild_test$correct, xlab="Mean score in Test", xlim=c(0,1))
hist(ripchild_transfer$correct, xlab="Mean score in Transfer", xlim=c(0,1))
hist(ripchild_individual$correct, xlab="Overall score (Test and Transfer)", xlim=c(0,1))
```

##Preparation of the data for running the GLMM

*Z-transformations for 'trial no' and 'age in months'
*Coding dummy variables for trialtype, sex and phase (order3)
*Centering the random slope components for these variables.

```{r include=FALSE}
#Preprocessing of data, Z-transforming trial no and age

ripchild$z.trialno<-as.vector(scale(ripchild$trialno))
ripchild$z.age=as.vector(scale(ripchild$agemonths))

ripchild$trialtype<-relevel(ripchild$trialtype, ref = "pen")
summary(ripchild)

```

```{r include=FALSE}
#Coding dummy variables before centering the slopes
ripchild$trialtype.sticker<-as.numeric(ripchild$trialtype==levels(ripchild$trialtype)[2])
ripchild$sex.m<-as.numeric(ripchild$sex==levels(ripchild$sex)[2])
ripchild$order3.transfer<-as.numeric(ripchild$order3==levels(ripchild$order3)[2])

#Centering the slopes
ripchild$trialtype.sticker.c<-ripchild$trialtype.sticker-mean(ripchild$trialtype.sticker)
ripchild$sex.m.c<-ripchild$sex.m -mean(ripchild$sex.m)
ripchild$order3.transfer.c<-ripchild$order3.transfer-mean(ripchild$order3.transfer)
```


```{r echo=FALSE}
source("./Roger_functions/diagnostic_fcns.r")
source("./Roger_functions/glmm_stability.r")
source("./Roger_functions/boot_glmm.r")
```



```{r include=FALSE}

# Guide on how to decide on the model...

xx.fe.re=fe.re.tab(fe.model="correct  ~
                  trialtype*order3+trialno+
                   z.age*order3+sex",
                   re="(1|id)",
                   data=data.frame(ripchild))
xx.fe.re$summary
ydata=xx.fe.re$data
```


##Test against chance
####Test and transfer phase

```{r, echo=TRUE}

##AS wants chance level performance to be reported first thing in the results section
testphase <-ripchild%>%
  filter(order3 == "test")
testchance <- glmer(correct ~ 1 + (z.trialno | id),data= testphase, family=binomial)
summary(testchance)

transferphase <-ripchild%>%
  filter(order3 == "transfer")
transferchance <- glmer(correct ~ 1 + (z.trialno | id),data= transferphase, family=binomial)
summary(transferchance)
```

**The intercept is significant: Performance in Test phase is above chance level.**



##Full model

```{r echo=TRUE}
contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))
full<-glmer(correct ~ trialtype*order3+z.trialno+z.age*order3+sex+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

The full model does not include box type (the location of the blue/pink box) as a random effect anymore due to convergence issues. But there is a singular fit warning here as you can see.

##Model assumptions

####Distribution of random effects
```{r echo=FALSE}
#An assumption of GLMMs is that the BLUPs are normally distributed. We use the following function to evaluate this. BLUPs are Best Linear Unbiased Predictors: the estimated grouping factor's level specific deviation from the common intercept or slope.
ranef.diagn.plot(full)
```

####Model stability

```{r echo=FALSE}
#summary reveals min and max of the estimates obtained after casewise deletion of levels of random effects together with the original estimates.
m.stab=glmm.model.stab(model.res=full, contr=contr)
m.stab.plot(m.stab$summary[,-1])
m.stab$summary
m.stab$detailed$warnings
```

####Multicollinearity

```{r echo=FALSE}
coll=lm(correct ~ trialtype+order3+z.trialno+z.age+sex, data=ripchild)
round(vif(coll),3)
#no vif issues
```

None of these seem to show a problem.

##Null model

```{r echo=TRUE}
null<-glmer(correct~ z.trialno+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

##Full and Null comparison
```{r echo=FALSE}
round(anova(null, full, test="Chisq"),3)
```

The full model is significantly different than the null model.

##Model output

####Coefficients

```{r echo=FALSE}
round(summary(full)$coefficients, 3)

```

####Individual predictor: Likelihood tests

```{r echo=FALSE}
xdrop1=drop1(full, test="Chisq",control=contr)
round(xdrop1,3)
```

####Confidence intervals for the full model with the interaction

```{r}
# model1CI=boot.glmm.pred(model.res=full, excl.warnings=T, nboots=1000, para=T)
# round(model1CI$ci.estimates, 3)
# save.image("model1_CIs.RData")
load("model1_CIs.RData")
round(model1CI$ci.estimates, 3)
```


**The interactions terms are not significant (although there is a trend for scores to vary by age) so I will remove them from the model.**

##Reduced model without the interactions
##Full model 2

```{r echo=TRUE}
full2=glmer(correct ~ trialtype+order3+z.trialno+z.age+sex+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

##Null model

```{r echo=TRUE}
null2<-glmer(correct~ z.trialno+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

####Full 2 and null comparision

```{r echo=FALSE}
round(anova(null2, full2, test="Chisq"),3)
```

Full model is significantly different than the null model.

##Model output
  
####Coefficients

```{r echo=FALSE}
round(summary(full2)$coefficients, 3)
```

####Individual predictors : Likelihod ratio tests
  
  
```{r echo=FALSE}
xdrop1=drop1(full2, test="Chisq",control=contr)
round(xdrop1,3)
```

**Age emerges as a significant predictor of scores. There also seems to be a pattern of scores varying by trial type.**

```{r include=FALSE}
# this is for me to check the variance added by random effects, number of observations and subjects etc.
print(summary(full2), corr=FALSE)
```

####Confidence intervals for the reduced model without the interaction

```{r}
# model2CI=boot.glmm.pred(model.res=full2, excl.warnings=T, nboots=1000, para=T)
# round(model2CI$ci.estimates, 3)
# save.image("model2_CIs.RData")
load("model2_CIs.RData")
round(model2CI$ci.estimates, 3)
```

####Calculating effect sizes for the final model

```{r echo=FALSE}
#install.packages("MuMIn")
library("MuMIn")
r.squaredGLMM(object=full2)

#as I’m not very passionate about effect sizes, I don’t go into detail here. the (marginal and conditional) effect sizes in the row headed delta are the ones according to a recent paper (Nakagawa et al., 2017) which you should report. Marginal R-squared is the variance explained by the fixed effects. The conditional R-squared is the variance explained by the fixed and random effects (the model)
```
Here we look at the row headed delta. Marginal R-squared is the variance explained by the fixed effects and the conditional R-squared is the variance explained by the fixed and random effects (the model).

##Scatterplot 1 to see the effect of age in the overall score. 

*Running a separate model for the CI calculation.

```{r echo=TRUE}
#running a new model for CI predictions
full.CI=glmer(correct ~ z.trialno+order3.transfer.c+trialtype.sticker.c+z.age+sex.m.c+(1+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)
```

*Getting confidence intervals for fitted values

```{r include=FALSE}
#boot.res=boot.glmm.pred(model.res=full.CI, excl.warnings=T, nboots=1000, resol=68, level=0.95, use="z.age", n.cores="all-1", para=T)#with para=T your computer will get very slow so I recommend using this option particularly when you don't need to work with the PC in parallel (the CI calculation will be a lot faster though)

#boot.res$ci.estimates
#boot.res$ci.predicted
#save.image("ripped_foil_child_CIs.RData")
load("ripped_foil_child_CIs.RData")
```

####Overall performance across age.

```{r echo=FALSE}
#there are 68 different data points (agemonths) and the argument resol of the function boot.glmm.pred was set to this value
#create sequence with agemonths against which lines should be plotted:


ripchild_individual2 <- ripchild %>%
  group_by(id,  z.age) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(z.age)%>%
    add_count(correct)

plot.xvals=seq(from=min(ripchild_individual$agemonths),to=max(ripchild_individual$agemonths),length.out=68)

##using ggplot but using the values from ripchild_individual instead)

scatter<-ggplot(ripchild_individual2, aes(x=plot.xvals, y=boot.res$ci.predicted$fitted)) + geom_point(aes(x=z.age, y=correct), size=ripchild_individual2$n, alpha=0.8)+
geom_ribbon(data=boot.res$ci.predicted, aes(x=z.age, ymin=boot.res$ci.predicted$lower.cl,ymax=boot.res$ci.predicted$upper.cl), fill="grey", alpha=0.5)+
geom_line(data=boot.res$ci.predicted, aes(x = z.age, y=boot.res$ci.predicted$fitted), lty=2)+
theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ xlab("Age in months")+ ylab("Mean correct responses")+ 
  ylim(0,1)+
  geom_hline(yintercept=0.50, color='red', linetype="dashed")

scatter

```

**The plot is not very informative as age in z-transformed. I correct this below.**

```{r echo=FALSE}
ripchild_individual <- ripchild %>%
  group_by(id, age, agemonths) %>% 
  summarize(correct = mean(correct))%>%
  ungroup()%>%
  group_by(agemonths)%>%
    add_count(correct)

plot.xvals=seq(from=min(ripchild_individual$agemonths),to=max(ripchild_individual$agemonths),length.out=68)

##using ggplot but using the values from ripchild_individual instead)

scatter<-ggplot(ripchild_individual, aes(x=plot.xvals, y=boot.res$ci.predicted$fitted)) + geom_point(aes(x=agemonths, y=correct), size=ripchild_individual$n, alpha=0.8)+
geom_ribbon(data=boot.res$ci.predicted, aes(x=plot.xvals, ymin=boot.res$ci.predicted$lower.cl,ymax=boot.res$ci.predicted$upper.cl), fill="grey", alpha=0.5)+
geom_line(data=boot.res$ci.predicted, aes(x = plot.xvals, y=boot.res$ci.predicted$fitted), lty=2)+ scale_x_continuous(breaks = seq(from = 36, to = 72, by = 6)) +  geom_segment(aes(x = 42, y = 0, xend = 42, yend = 0.5), arrow=arrow(length = unit(0.2, "cm")), colour="red")+
theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ xlab("Age in months")+ ylab("Mean correct responses")+ 
  ylim(0,1)+
  geom_hline(yintercept=0.50, color='red', linetype="dashed")

scatter

```

**Performance of younger children (up until 50 months of age) seem to be at or below chance level in this task. Older children seem to be doing better although there is a lot of variation (some are performing at ceiling level while some others are below chance).**

```{r include=FALSE}
# From Eva - need to put these top three lines, otherwise get an error msg.
par(mar=c(3, 3, 0.2, 0.2), mgp=c(1.7, 0.3, 0), las=1, tcl=-0.15)
plot(x=ripchild_individual$agemonths, y=ripchild_individual$correct, pch=19, las=2, ylim=c(0, 1),
xlab="Age in months", ylab="Mean correct responses")

plot.xvals=seq(from=min(ripchild_individual$agemonths), to=max(ripchild_individual$agemonths),length.out=68)

lines(x=plot.xvals, y=boot.res$ci.predicted$fitted,
lty=2)
lines(x=plot.xvals, y=boot.res$ci.predicted$lower.cl,
lty=3)
lines(x=plot.xvals, y=boot.res$ci.predicted$upper.cl,
lty=3)
abline(h = 0.5, col="red", lwd=3, lty=2)
```

####Running a new model for calculating CIs for the interaction model so we can plot the performance in test and transfer across ages separately.

```{r include=TRUE}
full.CI.int<-glmer(correct ~ trialtype.sticker.c+z.age*order3+sex.m.c+z.trialno+(1|id)+(0+z.trialno+order3.transfer.c+trialtype.sticker.c|id),data=ripchild, family=binomial, control=contr)

```


```{r include=FALSE}
#boot.res.int=boot.glmm.pred(model.res=full.CI.int, excl.warnings=T, nboots=1000, resol=136, level=0.95, use="z.age", n.cores="all-1", para=T) #resol=136 this time because data is aggregated for test and transfer separately.
#boot.res.int$ci.estimates
#boot.res.int$ci.predicted

#save.image("ripped_foil_child_CIs_interaction.RData")
load("ripped_foil_child_CIs_interaction.RData")
```

####Performance in Test and Transfer across age.

```{r echo=FALSE}
ripchild_separate <- ripchild %>%
  group_by(id,  z.age, order3) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(z.age, order3)%>%
    add_count(correct)

scatter.int<-ggplot(ripchild_separate, aes(x=boot.res.int$ci.predicted$z.age, y=boot.res.int$ci.predicted$fitted)) +
 geom_point(aes(x=z.age, y=correct), size=ripchild_separate$n, alpha=0.8) +    facet_wrap(~order3)+
geom_ribbon(data=boot.res.int$ci.predicted, aes(y=boot.res.int$ci.predicted$fitted, ymin=boot.res.int$ci.predicted$lower.cl, ymax=boot.res.int$ci.predicted$upper.cl), fill="grey", alpha=0.5)+geom_line(data=boot.res.int$ci.predicted, aes(y=boot.res.int$ci.predicted$fitted), lty=2)+
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ xlab("Age in months")+ ylab("Mean correct responses")+ 
  ylim(0,1)+
  geom_hline(yintercept=0.50, color='red', linetype="dashed")
  
scatter.int
  

```

**However, the age is z-transformed so the plot is not very informative. We change z-age to age in months.**

```{r echo=TRUE}
#do the same scatter plot for interaction with age not z-tranformed so the graph is more informative.
ripchild_separate2 <- ripchild %>%
  group_by(id,  agemonths, order3) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(agemonths, order3)%>%
    add_count(correct)

plot.int=seq(from=min(ripchild_separate2$agemonths),
             to=max(ripchild_separate2$agemonths),length.out=68)

ci.data<-boot.res.int$ci.predicted #I reduced the number of values for the confidence intervals. did not match the actual data.
toDelete<- seq(1, nrow(ci.data),2)
ci.data<-ci.data[toDelete,]

scatter.int2 <- ggplot(ci.data, aes(y = fitted)) +
  geom_point(data=ripchild_separate2, aes(x = agemonths, y = correct), size = ripchild_separate2$n, alpha = 0.8) +
  facet_wrap(~order3) + geom_segment(mapping=aes(x=44, xend=44, y=0, yend=0.5, order3="transfer"), arrow=arrow(length=unit(0.2,"cm")))+ geom_segment(mapping=aes(x=40, xend=40, y=0, yend=0.5, order3="test"), arrow=arrow(length=unit(0.2,"cm"))) +
  geom_ribbon(data = ci.data%>%filter(order3=="test"), aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl, group = order3), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data%>%filter(order3=="test"), aes(x = plot.int, y = fitted), lty = 2) +
    geom_ribbon(data = ci.data%>%filter(order3=="transfer"), aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl, group = order3), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data%>%filter(order3=="transfer"), aes(x = plot.int, y = fitted), lty = 2) + scale_x_continuous(breaks = seq(from = 36, to = 72, by = 4)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Age in months") +
  ylab("Mean correct responses") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.int2
```

**Younger children's (up until 50 months of age) scores in both test and transfer do not seem to differ from chance level compared to older children. The slope for the Transfer phase is also steeper than in Test level  meaning that the difference between the younger and older children's performance in Transfer is larger.**
This is great to see but we will leave this to the supplementary materials.


##First trial performance in Transfer phase
We expect younger children to make more errors when they switch from test to transfer if they are solving the task using an arbitrary rule.
*Getting the data for the first trial of Transfer

```{r echo=FALSE}
ripchild_1st_trial <- ripchild %>%
  filter(order3=="transfer" & trialno=="1") %>%
  group_by(id, age, agemonths, sex, trialtype, z.age, trialtype.sticker.c) %>% 
  summarize(correct)
ripchild_1st_trial$z.age=as.vector(scale(ripchild_1st_trial$agemonths))
str(ripchild_1st_trial)
```

####Performance in trial 1 of Transfer phase

```{r echo=TRUE}
##binomial test for the first trial performance in Transfer (CV asked for this and Amanda wants to stress the overall performance in Transfer trial 1)
binom.test(x=sum(ripchild_1st_trial$correct), n=length(ripchild_1st_trial$correct), p=0.5, alternative = "two.sided")

firsttrial <- glm(correct~z.age,data=ripchild_1st_trial,family=binomial(link = "logit"))
nullfirsttrial<-glm(correct~1, data=ripchild_1st_trial, family=binomial(link="logit"))
anova(nullfirsttrial, firsttrial, test="Chisq")
summary(firsttrial)$coefficients
confint(object=firsttrial)
cbind(coefficients(firsttrial), confint(object=firsttrial))
xdrop1=drop1(firsttrial, test="Chisq",control=contr)
round(xdrop1,3)
```

**Age emerges as a significant predictor of first trial performance of Transfer.**
We can plot this as:

```{r echo=FALSE}
#how about ggplot
ggplot(ripchild_1st_trial, aes(x=agemonths, y=correct)) + geom_point() + 
  stat_smooth(method="glm", fullrange = TRUE, lty=2, size=0.5, colour="Black", alpha=0.2,  method.args=list(family="binomial")) + scale_x_continuous(breaks = seq(from = 36, to = 72, by = 2)) +  geom_segment(aes(x = 50, y = 0, xend = 50, yend = 0.5), arrow=arrow(length = unit(0.2, "cm")), colour="red")+ylab("First trial performance of Transfer Phase")+xlab("Age in months")+theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + geom_hline(yintercept=0.50, color='red', linetype="dashed")

```

**Younger children tend to make more errors than older children**

```{r}
openended<-read.csv("open_ended_ripped.csv", header=T)
openended$z.age=as.vector(scale(openended$agemonths))
str(openended)
table(openended$explanation)
oequestion <- glm(explanation~z.age, data=openended, family=binomial(link = "logit"))
summary(oequestion)
anova(oequestion, test="Chisq")

#analysis of the open ended questions again after CV comments
oequestion <- glm(explanation~z.age, data=openended, family=binomial(link = "logit"))
nullopen <- glm(explanation ~ 1, data=openended, family=binomial(link="logit"))
anova(nullopen, oequestion, test="Chisq")
summary(oequestion)$coefficients
confint(object=oequestion)
cbind(coefficients(oequestion), confint(object=oequestion))
```

```{r}
ggplot(openended, aes(x=agemonths, y=explanation)) + geom_point() + 
  stat_smooth(method="glm", fullrange = TRUE, lty=2, size=0.5, colour="Black", alpha=0.2,  method.args=list(family="binomial"))+ scale_x_continuous(breaks = seq(from = 36, to = 72, by = 4))+ylab("Incorrect (0)/correct (1) explanations")+xlab("Age in months")+theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 


```


```{r}
#Comparing the performance of those who gave correct explanations to an age matched group
agematched<-read.csv("agematched_explanation.csv", header=T)
str(agematched)
table(agematched$explanation)



```














