---
title: "Ripped Foil - Monkeys"
author: "Zeynep Civelek"
date: "June 30, 2020"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, include=TRUE)

#This sets general options for the code chunks in the R Markdown file. The echo, warning, and message = FALSE hide the code chunks, warning messages, and other messages, where the ‘include=true’ will make all figures appear in the text. You can set some of these variables to TRUE, and hit Knit to see what they change. 
```

```{r, include=FALSE}
#PREPARE
R.Version()#for referencing, shows you which R version you are using
rm(list=ls())#removes any other items in your workspace
ls()#check whether workspace is empty
```

```{r, include=FALSE}
#Directory is set, data is added and all the libraries necessary for running the analyses are loaded.

ripmon1<-read.csv("ripped_monkey_test16.csv", header=T)
#setwd("C:/Users/zeyne/OneDrive/Desktop/ripped-foil/Ripped Foil")
# ripchild<-read.csv("ripped_child.csv", header=T)
#all the libraries necessary to run the analyses
library(lme4)
library(readr)
library(tidyverse)
library(sjPlot)
#install.packages('TMB', type='source')
library(ggthemes)
library(gridExtra)
library(reshape2)
library(car)
#library(psych)
library("ggpubr")
library(dplyr)

```
##Exploring data 

The structure of the data is shown below with all the variable names and levels of measurement.

```{r, echo=FALSE}
ripmon1$id<-as.factor(ripmon1$id)
ripmon1$sex<-as.factor(ripmon1$sex)
ripmon1$trialtype<-as.factor(ripmon1$trialtype)
ripmon1$phase2<-as.factor(ripmon1$phase2)
ripmon1$phase<-as.factor(ripmon1$phase)

str(ripmon1)
summary(ripmon1)
sd(ripmon1$trialno)
sd(ripmon1$age)

```

Exploring the warm-up 1 (locating the reward when there is no foil covering the cups) and warm-up 2 phases (locating the reward when the cups are covered with foil).
```{r, echo=FALSE}
ripmon_complete<-read.csv("ripped_monkey_complete.csv", header=T)
ripmon_warm1 <- ripmon_complete %>%
  filter(sessiontype=="warmup") %>%
  group_by(id, sex, age, sessionno, trialno) %>% 
  summarize(correct)

ripmon_warm2 <- ripmon_complete %>%
  filter(sessiontype=="foil") %>%
  group_by(id, sex, age, sessionno, trialno) %>% 
  summarize(correct)

range(ripmon_warm1$sessionno)
range(ripmon_warm2$sessionno)
```
**In warm-up 1, monkeys got up to 18 sessions to reach criterion and in warm-up 2, up to 6 sessions (or they'd receive up to 10 sessions). The minimum number of sessions to reach criterion was 2 (14/16 correct)**

##Histograms for Warm-up 1 and Warm-up 2 Phases
```{r, echo=FALSE}
agg_warm1 <- aggregate(ripmon_warm1$correct, by = list(id = ripmon_warm1$id), function(x) c(mean = mean(x), sd = sd(x), n = length(x)))
agg_warm1 <- do.call(data.frame, agg_warm1)
agg_warm1$se <- agg_warm1$x.sd / sqrt(agg_warm1$x.n)

colnames(agg_warm1) <- c("id", "mean", "sd", "n", "se")

agg_warm1$names <- c(paste(agg_warm1$id, "id"))

limits <- aes(ymax = agg_warm1$mean + agg_warm1$se,
              ymin = agg_warm1$mean - agg_warm1$se)

hist(agg_warm1$mean, xlab="Mean score in Warm-up 1", xlim=c(0,1))

agg_warm2 <- aggregate(ripmon_warm2$correct, by = list(id = ripmon_warm2$id), function(x) c(mean = mean(x), sd = sd(x), n = length(x)))
agg_warm2 <- do.call(data.frame, agg_warm2)
agg_warm2$se <- agg_warm2$x.sd / sqrt(agg_warm2$x.n)

colnames(agg_warm2) <- c("id", "mean", "sd", "n", "se")

agg_warm2$names <- c(paste(agg_warm2$id, "id"))

limits <- aes(ymax = agg_warm2$mean + agg_warm2$se,
              ymin = agg_warm2$mean - agg_warm2$se)

hist(agg_warm2$mean, xlab="Mean score in Warm-up 2", xlim=c(0,1))

```

I then aggregated the trial-by-trial data to create one score per monkey for Test and Transfer phases so I can see the distribution of scores.

```{r, echo=FALSE}
ripmon_individual <- ripmon1 %>%
  filter(!is.na(correct)) %>%
  group_by(id) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

ripmon_test <- ripmon1 %>%
   filter(!is.na(correct)) %>%
  filter(phase=="test") %>%
  group_by(id) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

ripmon_transfer <- ripmon1 %>%
   filter(!is.na(correct)) %>%
  filter(phase=="transfer") %>%
  group_by(id) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

ripmon_separate <- ripmon1 %>%
   filter(!is.na(correct)) %>%
  group_by(id, phase,age) %>% 
  summarize(correct = mean(correct))%>%
  add_count(correct)

```

##Histograms for Test and Transfer Phases and the overall score

```{r echo=FALSE}
hist(ripmon_test$correct, xlab="Mean score in Test", xlim=c(0,1))
hist(ripmon_transfer$correct, xlab="Mean score in Transfer", xlim=c(0,1))
hist(ripmon_individual$correct, xlab="Overall score (Test and Transfer)", xlim=c(0,1))
```

```{r, include=FALSE}
boxplot1 <- ggplot(ripmon_separate, aes(x=phase, y=correct)) + 
    geom_dotplot(binaxis='y', stackdir='center', dotsize=0.5) + ylim(0,1)
boxplot1 + stat_summary(fun.data="mean_sdl", fun.args = list(mult=1), geom="crossbar", width=0.5) + stat_summary(fun=mean, geom="point", color="red")+theme_bw() + stat_summary(fun.data=mean_sdl, fun.args = list(mult=1), 
                 geom="pointrange", color="red")+ theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+geom_hline(yintercept=0.50, color='red', linetype="dashed")+xlab("Phase") + ylab("Mean number of correct responses in 16 trials")
boxplot1
```

##Preparation of the data for running the GLMM

* Scaling age and trial number
* Coding categorical variables as dummy variables 
* Centering the slopes.

```{r, echo=FALSE}
#We scale variables so that the differences in the range of the variables do not influence the parameter estimations.
ripmon1$z.trialno<-as.vector(scale(ripmon1$trialno))
ripmon1$z.age=as.vector(scale(ripmon1$age))
ripmon1$trialtype<-relevel(ripmon1$trialtype, ref = "stick")

#coding dummy variables before centering the slopes
ripmon1$trialtype.food<-as.numeric(ripmon1$trialtype==levels(ripmon1$trialtype)[2])
ripmon1$sex.m<-as.numeric(ripmon1$sex==levels(ripmon1$sex)[2])
ripmon1$phase.transfer=as.numeric(ripmon1$phase==levels(ripmon1$phase)[2])

#centering the slopes: p-values of the factors can be influenced by the choice of reference category.
#by centering the factor for the random slope components the p-values should be the same irrespective of the choice of the reference level
ripmon1$trialtype.food.c<-ripmon1$trialtype.food-mean(ripmon1$trialtype.food)
ripmon1$phase.transfer.c=ripmon1$phase.transfer-mean(ripmon1$phase.transfer)
ripmon1$sex.m.c<-ripmon1$sex.m -mean(ripmon1$sex.m)
summary(ripmon1)

```

```{r, echo=FALSE}
source("./Roger_functions/diagnostic_fcns.r")
source("./Roger_functions/glmm_stability.r")
source("./Roger_functions/boot_glmm.r")
```

```{r, include=FALSE}

# Guide on how to decide on the model...

xx.fe.re=fe.re.tab(fe.model="correct  ~
                  trialtype*phase+trialno+
                   z.age+sex",
                   re="(1|id)",
                   data=data.frame(ripmon1))
xx.fe.re$summary
ydata=xx.fe.re$data
```

##Full model 

```{r, echo=TRUE}
contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))
full=glmer(correct ~ trialtype*phase+z.trialno+z.age+sex+(1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
```

The full model does not include box type (the location of the blue/pink box) as a random effect anymore due to convergence issues. The correlations between random slope and random intercept are not included either.

##Model assumptions

####Distribution of random effects
```{r, echo=FALSE}
#An assumption of GLMMs is that the BLUPs are normally distributed. We use the following function to evaluate this. BLUPs are Best Linear Unbiased Predictors: the estimated grouping factor's level specific deviation from the common intercept or slope.
ranef.diagn.plot(full)
```

####Model stability

```{r, echo=FALSE}
#summary reveals min and max of the estimates obtained after casewise deletion of levels of random effects together with the original estimates.
# m.stab=glmm.model.stab(model.res=full, contr=contr)
# m.stab.plot(m.stab$summary[,-1])
# m.stab$summary
# m.stab$detailed$warnings
```

####Multicollinearity

```{r, echo=FALSE}
coll=lm(correct ~ trialtype+phase+z.trialno+z.age+sex, data=ripmon1)
round(vif(coll),3)
#no vif issues
```

##Null model

```{r, echo=TRUE}
null=glmer(correct~z.trialno+(1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
```
**Added trial no to the null model to control for its effects.**

##Full and Null comparison
```{r, echo=FALSE}
round(anova(null, full, test="Chisq"),3)
```
**The full model is significantly different than the null model.**

##Model output

####Coefficients

```{r, echo=FALSE}
round(summary(full)$coefficients, 3)
```

####Individual predictor: Likelihood tests

```{r, echo=FALSE}
xdrop1=drop1(full, test="Chisq",control=contr)
round(xdrop1,3)
```
**There is a significant effect of sex. The interaction term is not significant.**


```{r, include=FALSE}
####Confidence intervals for the full model with the interaction

#ripmon_model1=boot.glmm.pred(model.res=full, excl.warnings=T, nboots=1000, para=T)
#save.image("ripmon_model1_CIs.RData")
load("ripmon_model1_CIs.RData")
round(ripmon_model1$ci.estimates, 3)

```

**I will remove the non-significant interaction term to reduce the model.**

##Reduced model without the interaction
##Full model 2

```{r, echo=TRUE}
full2=glmer(correct ~ trialtype+phase+z.trialno+z.age+sex+(1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
summary(full2)
```

I get the following warning: convergence code: 0, boundary (singular) fit: see ?isSingular.

**The model summary show that there is a significant effect of sex and trial type.**

##Null model 2

```{r, echo=TRUE}
null2<-glmer(correct~ z.trialno + (1|id)+(0+z.trialno+phase.transfer.c+trialtype.food.c|id),data=ripmon1, family=binomial, control=contr)
```

##Full and null model (2) comparion

```{r, echo=FALSE}
round(anova(null2, full2, test="Chisq"),3)
```
**There is a significant difference between the full and the null model.**

##Model output

####Coefficients

```{r, echo=FALSE}
round(summary(full2)$coefficients, 3)

```

####Individual predictors : Likelihood ratio tests
  
```{r echo=FALSE}
xdrop1=drop1(full2, test="Chisq",control=contr)
round(xdrop1,3)
```
**The trial type and sex are significant predictors of performance.**

```{r, include=FALSE}
# this is for me to check the variance added by random effects, number of observations and subjects etc.
print(summary(full2), corr=FALSE)
```


```{r, include=FALSE}
####Confidence intervals for the reduced model without the interaction

#ripmon_model2=boot.glmm.pred(model.res=full2, excl.warnings=T, nboots=1000, para=T)
#save.image("ripmon_model2_CIs.RData")
load("ripmon_model2_CIs.RData")
round(ripmon_model2$ci.estimates, 3)
```

####Calculating effect sizes for the final model

```{r, echo=FALSE}
#install.packages("MuMIn")
library("MuMIn")
r.squaredGLMM(object=full2)

#the (marginal and conditional) effect sizes in the row headed delta are the ones according to a recent paper (Nakagawa et al., 2017) which you should report. Marginal R-squared is the variance explained by the fixed effects. The conditional R-squared is the variance explained by the fixed and random effects (the model)
```
Here we look at the row headed delta. Marginal R-squared is the variance explained by the fixed effects and the conditional R-squared is the variance explained by the fixed and random effects (the model).


##Plotting the findings
```{r, echo=FALSE}
ripmon_individual2<-ripmon1 %>% 
  group_by(phase, trialtype, id) %>% 
  summarize(foil.correct = mean(correct)) %>% 
  add_count(foil.correct)

p1 <- ggplot(data=ripmon_individual2, aes(x=phase, y=foil.correct, group=phase)) + geom_boxplot()+geom_point(size = ripmon_individual2$n, colour = "darkgrey", alpha=0.3) + geom_line(aes(group = id), colour = "darkgrey", alpha = 0.5)+labs(x="",y="Mean number of correct choices")+theme_few()+ggtitle("Phase")+geom_hline(yintercept=0.5, linetype="dashed", color = "red")+ylim(0,1)
p1

#added p3 to show trial type in different colors
p4 <- ggplot(data=ripmon_individual2) + geom_boxplot(aes(x=phase, y=foil.correct, group=phase))+ylim(0,1)+geom_point(aes(x=phase, y=foil.correct, color=trialtype),size = ripmon_individual2$n, colour = "darkgrey", alpha=0.3) + geom_line(aes(x=phase, y= foil.correct, group=id, color=trialtype)) + labs(x="",y="Mean number of correct choices")+theme_few()+ggtitle("Phase")+geom_hline(yintercept=0.5, linetype="dashed", color = "red")
p4

ripmon_individual3 <- ripmon1 %>% 
  group_by(trialtype, id) %>% 
  summarize(foil.correct = mean(correct)) %>% 
  add_count(foil.correct)

p2 <- ggplot(data=ripmon_individual3, aes(x=trialtype, y=foil.correct, group=trialtype)) +geom_boxplot()+geom_line(aes(group = id), colour = "darkgrey", alpha = 0.5)+geom_point(size = ripmon_individual3$n, colour = "darkgrey", alpha=0.3) +geom_line(lty=2)+labs(x="",y="Mean number of correct choices")+ theme_few()+ggtitle("Trial type")+geom_hline(yintercept=0.5, linetype="dashed", color = "red")+ylim(0,1)

ripmon_individual4<-ripmon1 %>% 
  group_by(sex, id) %>% 
  summarize(foil.correct = mean(correct)) %>% 
  add_count(foil.correct)

p3 <- ggplot(data=ripmon_individual4, aes(x=sex, y=foil.correct, group=sex)) + geom_boxplot()+ylim(0,1)+ geom_point(size = ripmon_individual4$n, colour = "darkgrey", alpha=0.3) +labs(x="",y="Mean number of correct choices")+theme_few()+ggtitle("Sex")+geom_hline(yintercept=0.5, linetype="dashed", color = "red")

grid.arrange(p1, p2, nrow = 1, heights=unit(100, "mm"))
grid.arrange(p4, p2, nrow = 1, heights=unit(100, "mm"))

```
We will probably not include sex in the paper and it'll just be phase and trial type. Overall, monkeys performed better in food-stick trials than in stick-food trials and males better than females. Performance in Test and Transfer Phases did not differ.

```{r, include=FALSE}
#Doing a bar graph

ripaggregate <- aggregate(ripmon1$correct, by = list(order = ripmon1$order, phase = ripmon1$phase), function(x) c(mean = mean(x), sd = sd(x), n = length(x)))
ripaggregate <- do.call(data.frame, ripaggregate)
ripaggregate$se <- ripaggregate$x.sd / sqrt(ripaggregate$x.n)

colnames(ripaggregate) <- c("order", "phase", "mean", "sd", "n", "se")

ripaggregate$names <- c(paste(ripaggregate$order, "order",
                              ripaggregate$phase, "phase"))

limits <- aes(ymax = ripaggregate$mean + ripaggregate$se,
              ymin = ripaggregate$mean - ripaggregate$se)
p <- ggplot(data = ripaggregate, aes(x = factor(order), y = mean, fill = factor(phase)))
p + geom_bar(stat = "identity", position = position_dodge(0.9), width = 0.8) + geom_errorbar(limits, position = position_dodge(0.9), width = 0.10) + labs(x= "", y = "Mean correct responses") + ggtitle("Performance in test and transfer phases by trial type") + scale_fill_grey(name = "Phase") + geom_hline(yintercept=0.50, linetype="dashed", color="red", size=1)  + theme(legend.text = element_text(size = 8)) + ylim(0.00,1.00) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),panel.background = element_blank())

```

##Test against chance
####Test and transfer phase

```{r, echo=TRUE}
testphase <-ripmon1%>%
  filter(phase == "test")
testchance <- glmer(correct ~ 1 + (z.trialno | id),data= testphase, family=binomial)
summary(testchance)
confint(object=testchance)
```

**The intercept is significant: Performance in Test phase is above chance level.**

```{r, echo=TRUE}
transferphase <-ripmon1%>%
  filter(phase == "transfer")
transferchance <- glmer(correct ~ 1 + (z.trialno | id),data= transferphase, family=binomial)
summary(transferchance)
```

**Performanc in Transfer phase is above chance level as well. And there is no significant difference between the monkeys' performances in Test and Transfer phases.**

####Trial type

```{r, echo=FALSE}
sticktrials <-ripmon1%>%
  filter(trialtype == "stick")
stickchance <- glmer(correct ~ 1 + (z.trialno | id),data= sticktrials, family=binomial)
summary(stickchance)
```


```{r echo=FALSE}
foodtrials <-ripmon1%>%
  filter(trialtype == "food")
foodchance <- glmer(correct ~ 1 + (z.trialno | id),data= foodtrials, family=binomial)
summary(foodchance)
```
**Monkeys performed above chance levels in both stick-food and food-stick trials.**

##Correcting the p-values for the post-hoc tests using Hochberg correction (in the order: test, transfer, stick-food, food-stick)

```{r, echo=TRUE}
pvalues <- c(3.69e-06, 2.06e-08, 1.41e-05, 5.25e-09)
p.adjust(pvalues, method="hochberg")
```

##For comparison purposes with children, I run two additional tests
#### 1)First trial performance in Transfer phase: Is performance above chance?

```{r, echo=TRUE}
ripmon_1st_trial <- ripmon1 %>%
  filter(phase=="transfer" & sessionno=="1", trialno=="1")

sum(ripmon_1st_trial$correct)
length(ripmon_1st_trial$correct)

binom.test(x=sum(ripmon_1st_trial$correct), n=length(ripmon_1st_trial$correct), p=0.5, alternative = "two.sided")
```

**Question:** Is this the correct way of looking at chance performance here? Just in case, I'm including the t-test here as well.
**No, performance in the first trial of transfer phase does not differ from chance**

####2)Performance of monkeys in the first 8 trials of Test phase: Is it above chance level?

```{r, echo=FALSE}
ripmon_complete$z.trialno<-as.vector(scale(ripmon_complete$trialno))
ripmon_test8 <- ripmon_complete %>%
  filter(sessiontype=="test" & sessionno=="1")
test8chance<-glmer(correct ~ 1 + (z.trialno | id), data= ripmon_test8, family=binomial)
summary(test8chance)
```
**No, they are at chance level in the first 8 trials of Test phase.**

```{r}
#Amanda requested to see monkeys' performance throughout trials of Test and Transfer phase (faceted)- to show visually that they got many trials in Test phase. Christoph suggested doing 2 separate graphs for Test and Transfer to correct for the earlier CI issue in transfer phase- So I'll do that and then bind the graphs

###First do all the transformations 

ripmon3<-read.csv("ripped_monkey_complete_test.csv", header=T)
str(ripmon3)

ripmon3$id<-as.factor(ripmon3$id)
ripmon3$sex<-as.factor(ripmon3$sex)
ripmon3$phase<-as.factor(ripmon3$phase)
ripmon3$sessiontype<-as.factor(ripmon3$sessiontype)

ripmon3$trialtype<-as.factor(ripmon3$trialtype)


#We scale variables so that the differences in the range of the variables do not influence the parameter estimations.
ripmon3$z.trialno<-as.vector(scale(ripmon3$trialno))
ripmon3$z.age=as.vector(scale(ripmon3$age))
ripmon3$trialtype<-relevel(ripmon3$trialtype, ref = "stick")

#coding dummy variables before centering the slopes
ripmon3$trialtype.food<-as.numeric(ripmon3$trialtype==levels(ripmon3$trialtype)[2])
ripmon3$sex.m<-as.numeric(ripmon3$sex==levels(ripmon3$sex)[2])

#centering the slopes: p-values of the factors can be influenced by the choice of reference category.
#by centering the factor for the random slope components the p-values should be the same irrespective of the choice of the reference level
ripmon3$trialtype.food.c<-ripmon3$trialtype.food-mean(ripmon3$trialtype.food)
ripmon3$sex.m.c<-ripmon3$sex.m -mean(ripmon3$sex.m)

summary(ripmon3)

```




```{r}
##To do this I will run a new model for calculating CIs for the interaction model so we can plot the performance in test and transfer across trial numbers separately.

#contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))

#full.CI.int <- glmer(correct ~ trialtype.food.c + z.age + sex.m.c + z.trialno+ (1|id)+(0+z.trialno+trialtype.food.c|id), data=ripmon3, family=binomial, control=contr)

#boot.res.int=boot.glmm.pred(model.res=full.CI.int, excl.warnings=T, nboots=1000, resol=80, level=0.95, use="z.trialno", n.cores="all-1", para=T)

#boot.res.int$ci.estimates
#boot.res.int$ci.predicted

#save.image("ripped_mon_complete_testCIs.RData")
load("ripped_mon_complete_testCIs.RData")

```

```{r}
ripmon_separate2 <- ripmon3 %>%
  group_by(trialno) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(trialno)%>%
    add_count(correct)

plot.int=seq(from=min(ripmon_separate2$trialno),
             to=max(ripmon_separate2$trialno),length.out=80)

ci.data<-boot.res.int$ci.predicted #I reduced the number of values for the confidence intervals. did not match the actual data.


scatter.int2 <- ggplot(ci.data, aes(y = fitted)) +
  geom_point(data=ripmon_separate2, aes(x = trialno, y = correct), size = ripmon_separate2$n, alpha = 0.8) +
  geom_ribbon(data = ci.data, aes(x = plot.int, y = fitted, ymin = lower.cl, ymax = upper.cl), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data, aes(x = plot.int, y = fitted), lty = 2)+ scale_x_continuous(breaks = c(1,8,16,24,32,40,48,56,64,72,80)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Number of trials in Test") +
  ylab("Correct response") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.int2
```

```{r}
#Amanda requested to see monkeys' performance throughout trials of Test and Transfer phase (faceted)- to show visually that they got many trials in Test phase. Christoph suggested doing 2 separate graphs for Test and Transfer to correct for the earlier CI issue in transfer phase- So I'll do that and then bind the graphs

###First do all the transformations 

ripmon4<-read.csv("ripped_monkey_complete_transfer.csv", header=T)
str(ripmon4)

ripmon4$id<-as.factor(ripmon4$id)
ripmon4$sex<-as.factor(ripmon4$sex)
ripmon4$sessiontype<-as.factor(ripmon4$sessiontype)

ripmon4$trialtype<-as.factor(ripmon4$trialtype)


#We scale variables so that the differences in the range of the variables do not influence the parameter estimations.
ripmon4$z.trialno<-as.vector(scale(ripmon4$trialno))
ripmon4$trialtype<-relevel(ripmon4$trialtype, ref = "stick")
ripmon4$z.age=as.vector(scale(ripmon4$age))

#coding dummy variables before centering the slopes
ripmon4$trialtype.food<-as.numeric(ripmon4$trialtype==levels(ripmon4$trialtype)[2])
ripmon4$sex.m<-as.numeric(ripmon4$sex==levels(ripmon4$sex)[2])

#centering the slopes: p-values of the factors can be influenced by the choice of reference category.
#by centering the factor for the random slope components the p-values should be the same irrespective of the choice of the reference level
ripmon4$trialtype.food.c<-ripmon4$trialtype.food-mean(ripmon4$trialtype.food)
ripmon4$sex.m.c<-ripmon4$sex.m -mean(ripmon4$sex.m)

summary(ripmon4)

```

```{r}
##To do this I will run a new model for calculating CIs for the interaction model so we can plot the performance in test and transfer across trial numbers separately.

#contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))

#full.CI.int2 <- glmer(correct ~ trialtype.food.c + z.age + sex.m.c + z.trialno+ (1|id)+(0+z.trialno+trialtype.food.c|id), data=ripmon4, family=binomial, control=contr)

#boot.res.int2=boot.glmm.pred(model.res=full.CI.int2, excl.warnings=T, nboots=1000, resol=16, level=0.95, use="z.trialno", n.cores="all-1", para=T)

#boot.res.int2$ci.estimates
#boot.res.int2$ci.predicted

#save.image("ripped_mon_complete_transferCIs.RData")
load("ripped_mon_complete_transferCIs.RData")



```


```{r}
ripmon_separate3 <- ripmon4 %>%
  group_by(trialno) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(trialno)%>%
    add_count(correct)

plot.int2=seq(from=min(ripmon_separate3$trialno),
             to=max(ripmon_separate3$trialno),length.out=16)

ci.data2<-boot.res.int2$ci.predicted #I reduced the number of values for the confidence intervals. did not match the actual data.


scatter.int3 <- ggplot(ci.data2, aes(y = fitted)) +
  geom_point(data=ripmon_separate3, aes(x = trialno, y = correct), size = ripmon_separate3$n, alpha = 0.8) +
  geom_ribbon(data = ci.data2, aes(x = plot.int2, y = fitted, ymin = lower.cl, ymax = upper.cl), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data2, aes(x = plot.int2, y = fitted), lty = 2)+ scale_x_continuous(breaks =c(1,8,16)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Number of trials in Transfer") +
  ylab("Correct response") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.int3

ggarrange(scatter.int2, scatter.int3, widths= c(7,7), heights = c(10, 10),
          ncol = 2, nrow = 1)
ggsave("test transfer across trials exp 2.png")
```


```{r}
##Christoph said looking at these graphs: "also one can see that the glmm is not really good at fitting data like in the transfer phase. it does not capture at all the rapid improvement between the first and second trial. it is more suited for more gradual changes in performance. I'd be curious how the results would look like if you included (z.trial + I(z.trial^2)) instead of z.trial in the transfer model.  I'm not saying that we should use it for the paper but it would be interesting to see how the model would fit the data with the polynomial included

contr<-glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=10000000))

full.CI.transfer <- glmer(correct ~ trialtype.food.c + z.age + sex.m.c + z.trialno + I(z.trialno^2)+ (1|id)+(0+z.trialno+ I(z.trialno^2)+trialtype.food.c|id), data=ripmon4, family=binomial, control=contr)

boot.res.transfer=boot.glmm.pred(model.res=full.CI.transfer, excl.warnings=T, nboots=10, resol=16, level=0.95, use="z.trialno+I(z.trialno^2)", n.cores="all-1", para=T)

boot.res.transfer$ci.estimates
boot.res.transfer$ci.predicted

#save.image("ripped_mon_complete_transferCIs.RData")
load("ripped_mon_complete_transferCIs.RData")



```

```{r}
ripmon_separate4 <- ripmon4 %>%
  group_by(trialno) %>% 
  summarize(correct= mean(correct))%>%
  ungroup()%>%
  group_by(trialno)%>%
    add_count(correct)

plot.transfer=seq(from=min(ripmon_separate4$trialno),
             to=max(ripmon_separate4$trialno),length.out=16)

ci.data3<-boot.res.transfer$ci.predicted


scatter.transfer <- ggplot(ci.data3, aes(y = fitted)) +
  geom_point(data=ripmon_separate4, aes(x = trialno, y = correct), size = ripmon_separate4$n, alpha = 0.8) +
  geom_ribbon(data = ci.data3, aes(x = plot.transfer, y = fitted, ymin = lower.cl, ymax = upper.cl), fill = "grey", alpha = 0.5) +
  geom_line(data = ci.data3, aes(x = plot.transfer, y = fitted), lty = 2)+ scale_x_continuous(breaks =c(1,8,16)) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab("Number of trials in Transfer") +
  ylab("Correct response") +
  ylim(0, 1) +
  geom_hline(yintercept = 0.50, color = "red", linetype = "dashed")

scatter.transfer

ggarrange(scatter.int2, scatter.int3, widths= c(7,7), heights = c(10, 10),
          ncol = 2, nrow = 1)
```

